---
title: "BoarScript-FUKU"
output:
  html_document: default
  pdf_document: default
date: "2024-01-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Library Calls
```{r}
library(devtools)
library(readxl)
library(unmarked)
library(lme4)
library(arm)
library(MuMIn)
library(AICcmodavg)
library(ggplot2)
```

Loading and Formatting Data
```{r}
FukuBoar <- readxl::read_excel("/Users/sophiavrh/Desktop/complete data w approximations.xlsx", sheet = "Sheet1")
FukuBoar$EZ <- as.factor(FukuBoar$EZ)
y1 <- as.matrix(FukuBoar[,2:5])
obscov1 <- list("camdays"=FukuBoar[,6:9])
sitecov1 <- as.data.frame(FukuBoar[,10:21])

#create unmarked frame
UMFfuku <- unmarked::unmarkedFramePCount(y=y1, siteCovs = sitecov1, obsCovs = obscov1)
summary(UMFfuku)
hist(na.omit(UMFfuku@y),breaks=831,xlab="Number of Wild Boar", main="Distribution of Wild Boar sightings per Camera Year")

```

Null Model Generation:
```{r}
null1 <- unmarked::pcount(formula = ~camdays ~ 1, data = UMFfuku, K = 831, mixture = "NB")
MuMIn::AICc(null1)
#12350.28
```

Collinearity?:
```{r}
cor(FukuBoar$Radiation, FukuBoar$Lat)
cor(FukuBoar$Radiation, FukuBoar$Long) #0.685
cor(FukuBoar$Radiation, FukuBoar$Elevation)
cor(FukuBoar$Radiation, FukuBoar$Distance_Edge)
cor(FukuBoar$Radiation, FukuBoar$Distance_to_Highway)
cor(FukuBoar$Radiation, FukuBoar$Distance_Water)
cor(FukuBoar$Radiation, FukuBoar$NDVI)
cor(FukuBoar$Radiation, FukuBoar$Veg_Midstory)
cor(FukuBoar$Radiation, FukuBoar$Veg_Ground)
cor(FukuBoar$Radiation, FukuBoar$Canopy_Cover)

cor(FukuBoar$Elevation, FukuBoar$Lat)
cor(FukuBoar$Elevation, FukuBoar$Long) #-0.6875
cor(FukuBoar$Elevation, FukuBoar$Distance_Edge)
cor(FukuBoar$Elevation, FukuBoar$Distance_to_Highway)
cor(FukuBoar$Elevation, FukuBoar$Distance_Water) #0.5726
cor(FukuBoar$Elevation, FukuBoar$NDVI)
cor(FukuBoar$Elevation, FukuBoar$Veg_Midstory)
cor(FukuBoar$Elevation, FukuBoar$Veg_Ground)
cor(FukuBoar$Elevation, FukuBoar$Canopy_Cover)

cor(FukuBoar$Distance_Edge, FukuBoar$Lat)
cor(FukuBoar$Distance_Edge, FukuBoar$Long) 
cor(FukuBoar$Distance_Edge, FukuBoar$Distance_to_Highway)
cor(FukuBoar$Distance_Edge, FukuBoar$Distance_Water) 
cor(FukuBoar$Distance_Edge, FukuBoar$NDVI)
cor(FukuBoar$Distance_Edge, FukuBoar$Veg_Midstory)
cor(FukuBoar$Distance_Edge, FukuBoar$Veg_Ground)
cor(FukuBoar$Distance_Edge, FukuBoar$Canopy_Cover)

cor(FukuBoar$Distance_to_Highway, FukuBoar$Lat)
cor(FukuBoar$Distance_to_Highway, FukuBoar$Long) 
cor(FukuBoar$Distance_to_Highway, FukuBoar$Distance_Water) 
cor(FukuBoar$Distance_to_Highway, FukuBoar$NDVI)
cor(FukuBoar$Distance_to_Highway, FukuBoar$Veg_Midstory)
cor(FukuBoar$Distance_to_Highway, FukuBoar$Veg_Ground)
cor(FukuBoar$Distance_to_Highway, FukuBoar$Canopy_Cover)

cor(FukuBoar$Distance_Water, FukuBoar$Lat)
cor(FukuBoar$Distance_Water, FukuBoar$Long) 
cor(FukuBoar$Distance_Water, FukuBoar$NDVI)
cor(FukuBoar$Distance_Water, FukuBoar$Veg_Midstory)
cor(FukuBoar$Distance_Water, FukuBoar$Veg_Ground)
cor(FukuBoar$Distance_Water, FukuBoar$Canopy_Cover)

cor(FukuBoar$NDVI, FukuBoar$Lat)
cor(FukuBoar$NDVI, FukuBoar$Long) 
cor(FukuBoar$NDVI, FukuBoar$Veg_Midstory)
cor(FukuBoar$NDVI, FukuBoar$Veg_Ground)
cor(FukuBoar$NDVI, FukuBoar$Canopy_Cover)

cor(FukuBoar$Veg_Midstory, FukuBoar$Lat)
cor(FukuBoar$Veg_Midstory, FukuBoar$Long) 
cor(FukuBoar$Veg_Midstory, FukuBoar$Veg_Ground) #0.65325
cor(FukuBoar$Veg_Midstory, FukuBoar$Canopy_Cover)

cor(FukuBoar$Veg_Ground, FukuBoar$Lat)
cor(FukuBoar$Veg_Ground, FukuBoar$Long) 
cor(FukuBoar$Veg_Ground, FukuBoar$Canopy_Cover)

cor(FukuBoar$Canopy_Cover, FukuBoar$Lat)
cor(FukuBoar$Veg_Ground, FukuBoar$Long) 

FukuBoarcovs <- FukuBoar[, 10:21]
V <- cor(FukuBoarcovs)
library(corrplot)
corrplot(V, method = "number")

#longitude-elevation, longitude-radiation, vegmid-vegground
cor.test(FukuBoarcovs$Elevation, FukuBoarcovs$Long)
cor.test(FukuBoarcovs$Radiation, FukuBoarcovs$Long)
cor.test(FukuBoarcovs$Veg_Midstory, FukuBoarcovs$Veg_Ground)
```


Single Covariate Models:
Will be testing a zero-inflated poisson (zip) and negative binomial (nb) distribution
```{r}
nb01 <- unmarked::pcount(~camdays ~ Lat, data = UMFfuku, K = 831, mixture = "NB")
nb02 <- unmarked::pcount(~camdays ~ Long, data = UMFfuku, K = 831, mixture = "NB")
nb03 <- unmarked::pcount(~camdays ~ Elevation, data = UMFfuku, K = 831, mixture = "NB")
nb04 <- unmarked::pcount(~camdays ~ Radiation, data = UMFfuku, K = 831, mixture = "NB")
nb05 <- unmarked::pcount(~camdays ~ EZ, data = UMFfuku, K = 831, mixture = "NB")
nb06 <- unmarked::pcount(~camdays ~ Distance_Edge, data = UMFfuku, K = 831, mixture = "NB")
nb07 <- unmarked::pcount(~camdays ~ Distance_to_Highway, data = UMFfuku, K = 831, mixture = "NB")
nb08 <- unmarked::pcount(~camdays ~ Distance_Water, data = UMFfuku, K = 831, mixture = "NB")
nb09 <- unmarked::pcount(~camdays ~ NDVI, data = UMFfuku, K = 831, mixture = "NB")
nb010 <- unmarked::pcount(~camdays ~ Veg_Midstory, data = UMFfuku, K = 831, mixture = "NB")
nb011 <- unmarked::pcount(~camdays ~ Veg_Ground, data = UMFfuku, K = 831, mixture = "NB")
nb012 <- unmarked::pcount(~camdays ~ Canopy_Cover, data = UMFfuku, K = 831, mixture = "NB")
zip01 <- unmarked::pcount(~camdays ~ Lat, data = UMFfuku, K = 831, mixture = "ZIP")
zip02 <- unmarked::pcount(~camdays ~ Long, data = UMFfuku, K = 831, mixture = "ZIP")
zip03 <- unmarked::pcount(~camdays ~ Elevation, data = UMFfuku, K = 831, mixture = "ZIP")
zip04 <- unmarked::pcount(~camdays ~ Radiation, data = UMFfuku, K = 831, mixture = "ZIP")
zip05 <- unmarked::pcount(~camdays ~ EZ, data = UMFfuku, K = 831, mixture = "ZIP")
zip06 <- unmarked::pcount(~camdays ~ Distance_Edge, data = UMFfuku, K = 831, mixture = "ZIP")
zip07 <- unmarked::pcount(~camdays ~ Distance_to_Highway, data = UMFfuku, K = 831, mixture = "ZIP")
zip08 <- unmarked::pcount(~camdays ~ Distance_Water, data = UMFfuku, K = 831, mixture = "ZIP")
zip09 <- unmarked::pcount(~camdays ~ NDVI, data = UMFfuku, K = 831, mixture = "ZIP")
zip010 <- unmarked::pcount(~camdays ~ Veg_Midstory, data = UMFfuku, K = 831, mixture = "ZIP")
zip011 <- unmarked::pcount(~camdays ~ Veg_Ground, data = UMFfuku, K = 831, mixture = "ZIP")
zip012 <- unmarked::pcount(~camdays ~ Canopy_Cover, data = UMFfuku, K = 831, mixture = "ZIP")

#Testing model fit and returning AIC values
fitest1 <- unmarked::fitList(nb01, nb02, nb03, nb04, nb05, nb06, nb07, nb08, nb09, nb010, nb011, nb012, zip01, zip02, zip03, zip04, zip05, zip06, zip07, zip08, zip09, zip010, zip011, zip012)
testset1 <- unmarked::modSel(fitest1) 
AICtestset1 <- data.frame(testset1@Full[["model"]],testset1@Full[["AIC"]])
#negative binomial models for each covariate return the lowest AIC, so all future models will be made with the negative binomial mixture
testset1
```

Corrected AIC values:
```{r}
MuMIn::AICc(nb01, nb02, nb03, nb04, nb05, nb06, nb07, nb08, nb09, nb010, nb011, nb012)
#nb05 has the lowest AIC and AICc (12341.15)

newlist <- list(null1, nb01, nb02, nb03, nb04, nb05, nb06, nb07, nb08, nb09, nb010, nb011, nb012)
AICcmodavg::aictab(newlist, modnames = c("null1", "nb01", "nb02", "nb03", "nb04", "nb05", "nb06", "nb07", "nb08", "nb09", "nb010", "nb011", "nb012"))
```


Model Set #2: 
Additive models, all possible combinations EZ and another covariate
```{r}
fm.11 <- unmarked::pcount(~camdays ~ Lat + EZ, data = UMFfuku, K = 831, mixture = "NB")
fm.12 <- unmarked::pcount(~camdays ~ Long + EZ, data = UMFfuku, K = 831, mixture = "NB")
fm.13 <- unmarked::pcount(~camdays ~ Elevation + EZ, data = UMFfuku, K = 831, mixture = "NB")
fm.14 <- unmarked::pcount(~camdays ~ Radiation + EZ, data = UMFfuku, K = 831, mixture = "NB")
fm.15 <- unmarked::pcount(~camdays ~ EZ + Distance_Edge, data = UMFfuku, K = 831, mixture = "NB")
fm.16 <- unmarked::pcount(~camdays ~ EZ + Distance_to_Highway, data = UMFfuku, K = 831, mixture = "NB")
fm.17 <- unmarked::pcount(~camdays ~ EZ + Distance_Water, data = UMFfuku, K = 831, mixture = "NB")
fm.18 <- unmarked::pcount(~camdays ~ EZ + NDVI, data = UMFfuku, K = 831, mixture = "NB")
fm.19 <- unmarked::pcount(~camdays ~ EZ + Veg_Midstory, data = UMFfuku, K = 831, mixture = "NB")
fm.110 <- unmarked::pcount(~camdays ~ EZ + Veg_Ground, data = UMFfuku, K = 831, mixture = "NB")
fm.111 <- unmarked::pcount(~camdays ~ EZ + Canopy_Cover, data = UMFfuku, K = 831, mixture = "NB")
```

Corrected AIC:
```{r}
newlist3 <- list(null1, nb05, nb011, nb09, nb04, fm.11, fm.12, fm.13, fm.14, fm.15, fm.16, fm.17, fm.18, fm.19, fm.110, fm.111)

AICcmodavg::aictab(newlist3, modnames = c("null1", "nb05", "nb011", "nb09", "nb04", "fm.11", "fm.12", "fm.13", "fm.14", "fm.15", "fm.16", "fm.17", "fm.18", "fm.19", "fm.110", "fm.111"))
```

Models for Cameras INSIDE the Exclusion Zone Only:
```{r}
FukuBoar2 <- FukuBoar[FukuBoar$EZ==1, ]
y2 <- as.matrix(FukuBoar2[,2:5])
obscov2 <- list("camdays"=FukuBoar2[,6:9])
sitecov2 <- as.data.frame(FukuBoar2[,10:21])

#create unmarked frame
UMFfuku2 <- unmarked::unmarkedFramePCount(y=y2, siteCovs = sitecov2, obsCovs = obscov2)
summary(UMFfuku2)

#single variable models
nb2.1 <- unmarked::pcount(~camdays ~ Lat, data = UMFfuku2, K = 831, mixture = "NB")
nb2.2 <- unmarked::pcount(~camdays ~ Long, data = UMFfuku2, K = 831, mixture = "NB")
nb2.3 <- unmarked::pcount(~camdays ~ Elevation, data = UMFfuku2, K = 831, mixture = "NB")
nb2.4 <- unmarked::pcount(~camdays ~ Radiation, data = UMFfuku2, K = 831, mixture = "NB")
nb2.5 <- unmarked::pcount(~camdays ~ Distance_Edge, data = UMFfuku2, K = 831, mixture = "NB")
nb2.6 <- unmarked::pcount(~camdays ~ Distance_to_Highway, data = UMFfuku2, K = 831, mixture = "NB")
nb2.7 <- unmarked::pcount(~camdays ~ Distance_Water, data = UMFfuku2, K = 831, mixture = "NB")
nb2.8 <- unmarked::pcount(~camdays ~ NDVI, data = UMFfuku2, K = 831, mixture = "NB")
nb2.9 <- unmarked::pcount(~camdays ~ Veg_Midstory, data = UMFfuku2, K = 831, mixture = "NB")
nb2.10 <- unmarked::pcount(~camdays ~ Veg_Ground, data = UMFfuku2, K = 831, mixture = "NB")
nb2.11 <- unmarked::pcount(~camdays ~ Canopy_Cover, data = UMFfuku2, K = 831, mixture = "NB")

#Testing model fit and returning AIC values
fitest2.1 <- unmarked::fitList(nb2.1, nb2.2, nb2.3, nb2.4, nb2.5, nb2.6, nb2.7, nb2.8, nb2.9, nb2.10, nb2.11)
testset2.1 <- unmarked::modSel(fitest2.1) 
testset2.1
```

```{r}
AICc2.1 <- MuMIn::AICc(nb2.1, nb2.2, nb2.3, nb2.4, nb2.5, nb2.6, nb2.7, nb2.8, nb2.9, nb2.10, nb2.11)
#nb2.3 has the lowest AIC (10843.08) and AICc (10844.19)

null2 <- unmarked::pcount(formula = ~camdays ~ 1, data = UMFfuku2, K = 831, mixture = "NB")
  
newlist4 <- list(null2, nb2.1, nb2.2, nb2.3, nb2.4, nb2.5, nb2.6, nb2.7, nb2.8, nb2.9, nb2.10, nb2.11)
AICcmodavg::aictab(newlist4, modnames = c("null2", "nb2.1", "nb2.2", "nb2.3", "nb2.4", "nb2.5", "nb2.6", "nb2.7", "nb2.8", "nb2.9", "nb2.10", "nb2.11"))

```

Model Set 2, using Elevation for additive models
```{r}
fm2.1 <- unmarked::pcount(~camdays ~ Elevation + Lat, data = UMFfuku2, K = 831, mixture = "NB")
fm2.2 <- unmarked::pcount(~camdays ~ Elevation + Long, data = UMFfuku2, K = 831, mixture = "NB")
fm2.3 <- unmarked::pcount(~camdays ~ Elevation + Radiation, data = UMFfuku2, K = 831, mixture = "NB")
fm2.4 <- unmarked::pcount(~camdays ~ Elevation + Distance_Edge, data = UMFfuku2, K = 831, mixture = "NB")
fm2.5 <- unmarked::pcount(~camdays ~ Elevation + Distance_to_Highway, data = UMFfuku2, K = 831, mixture = "NB")
fm2.6 <- unmarked::pcount(~camdays ~ Elevation + Distance_Water, data = UMFfuku2, K = 831, mixture = "NB")
fm2.7 <- unmarked::pcount(~camdays ~ Elevation + NDVI, data = UMFfuku2, K = 831, mixture = "NB")
fm2.8 <- unmarked::pcount(~camdays ~ Elevation + Veg_Midstory, data = UMFfuku2, K = 831, mixture = "NB")
fm2.9 <- unmarked::pcount(~camdays ~ Elevation + Veg_Ground, data = UMFfuku2, K = 831, mixture = "NB")
fm2.10 <- unmarked::pcount(~camdays ~ Elevation + Canopy_Cover, data = UMFfuku2, K = 831, mixture = "NB")

#Testing model fit and returning AIC values
fitest2.2 <- unmarked::fitList(fm2.1, fm2.2, fm2.3, fm2.4, fm2.5, fm2.6, fm2.7, fm2.8, fm2.9, fm2.10)
testset2.2 <- unmarked::modSel(fitest2.2) 
testset2.2
```

```{r}
AICc2.2 <- MuMIn::AICc(fm2.1, fm2.2, fm2.3, fm2.4, fm2.5, fm2.6, fm2.7, fm2.8, fm2.9, fm2.10)
#fm2.7 has the lowest AIC (11127.81) and AICc (11129.97) but the original 1 variable model for elevation has a lower value for both
AICc2.2

newlist4 <- list(null2, nb2.3, fm2.1, fm2.2, fm2.3, fm2.4, fm2.5, fm2.6, fm2.7, fm2.8, fm2.9, fm2.10)
AICcmodavg::aictab(newlist4, modnames = c("null2", "nb2.3", "fm2.1", "fm2.2", "fm2.3", "fm2.4", "fm2.5", "fm2.6", "fm2.7", "fm2.8", "fm2.9", "fm2.10"))
```


RAI Barplot per Year
```{r}
FukuRAIs <- read_excel("complete data w approximations.xlsx", sheet = "RAI per year")
M2015 <- mean(na.omit(FukuRAIs$`2015`))
M2016 <- mean(na.omit(FukuRAIs$`2016`))
M2017 <- mean(na.omit(FukuRAIs$`2017`))
M2018 <- mean(na.omit(FukuRAIs$`2018`))

newframe <- data.frame(Year=c('2015', '2016', '2017', '2018'), value=c(M2015, M2016, M2017, M2018))

V2015 <- var(na.omit(FukuRAIs$`2015`))
V2016 <- var(na.omit(FukuRAIs$`2016`))
V2017 <- var(na.omit(FukuRAIs$`2017`))
V2018 <- var(na.omit(FukuRAIs$`2018`))

newframe2 <- data.frame(Year=c('2015', '2016', '2017', '2018'),  value=c(M2015, M2016, M2017, M2018), varr=c(V2015, V2016, V2017, V2018))

ggplot(data = newframe2, aes(x=Year, y=value)) + geom_bar(stat = "identity", fill =  c("#481567FF", "#2A788EFF", "#75D054FF", "#FDE725FF")) + ylab("Mean RAI Value") + geom_errorbar(aes(x=Year, ymin=value-varr, ymax=value+varr), width = 0.2, color = "black", alpha = 0.9) + theme_bw(base_size = 15)

```

Boxplot instead?
```{r}
boxframe15 <- data.frame(RAI = na.omit(FukuRAIs$`2015`), Year = rep(2015, 42))
boxframe16 <- data.frame(RAI = na.omit(FukuRAIs$`2016`), Year = rep(2016, 52))
boxframe17 <- data.frame(RAI = na.omit(FukuRAIs$`2017`), Year = rep(2017, 57))
boxframe18 <- data.frame(RAI = na.omit(FukuRAIs$`2018`), Year = rep(2018, 54))


boxframe <- rbind(boxframe15, boxframe16, boxframe17, boxframe18)

boxframe$Year <- as.factor(boxframe$Year)

ggplot(data = boxframe, aes(x=Year, y=RAI)) + geom_boxplot(fill = c("#481567FF", "#2A788EFF", "#75D054FF", "#FDE725FF")) + coord_flip(ylim=c(0,4)) + theme_bw(base_size = 15)
```

```{r}
#Log(x+1) transforming RAI
boxframe15$transform <- log(boxframe15$RAI + 1)
boxframe16$transform <- log(boxframe16$RAI + 1)
boxframe17$transform <- log(boxframe17$RAI + 1)
boxframe18$transform <- log(boxframe18$RAI + 1)

boxframe <- rbind(boxframe15, boxframe16, boxframe17, boxframe18)
boxframe$Year <- as.factor(boxframe$Year)

ggplot(data = boxframe, mapping = aes(x=Year, y=transform)) + geom_boxplot(fill = c("#481567FF", "#2A788EFF", "#75D054FF", "#FDE725FF")) + labs(y = "Log(RAI + 1)", title = "Fukushima") + theme_bw(base_size = 15)

```
