---
title: "BoarScript-Cher.Rmd"
output:
  html_document: default
  pdf_document: default
date: "2024-03-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Library Calls
```{r}
library(devtools)
library(readxl)
library(unmarked)
library(lme4)
library(arm)
library(MuMIn)
library(AICcmodavg)
library(ggplot2)
```

Loading and Formatting Data
```{r}
setwd("/Users/sophiavrh/Desktop/Mousseau Lab")

CherBoar <- readxl::read_excel("Sophia Thesis- Cher ALL.xlsx", sheet = "final R")
y1 <- as.matrix(CherBoar[,2:9])
obscov1 <- list("camdays"=CherBoar[,10:17])
sitecov1 <- as.data.frame(CherBoar[,18:29])

#create unmarked frame
UMFcher <- unmarked::unmarkedFramePCount(y=y1, siteCovs = sitecov1, obsCovs = obscov1)
summary(UMFcher)
```

Null Model Generation:
```{r}
null <- unmarked::pcount(formula = ~camdays ~ 1, data = UMFcher, K = 858, mixture = "NB")
MuMIn::AICc(null)
#4281.568
```

Collinearity?:
```{r}
cor(CherBoar$Lat., CherBoar$Sr90) #0.4315
cor(CherBoar$Lat., CherBoar$Cs137) #0.4597
cor(CherBoar$Lat., CherBoar$Am241) #0.42696
cor(CherBoar$Lat., CherBoar$Pu238) #0.43843
cor(CherBoar$Lat., CherBoar$Pu239) #0.4401
cor(CherBoar$Lat., CherBoar$Pu240) #0.4384

cor(CherBoar$Long., CherBoar$Sr90) #0.2157
cor(CherBoar$Long., CherBoar$Cs137) #0.1065
cor(CherBoar$Long., CherBoar$Am241) #0.2092
cor(CherBoar$Long., CherBoar$Pu238) #0.2159
cor(CherBoar$Long., CherBoar$Pu239) #0.2125
cor(CherBoar$Long., CherBoar$Pu240) #0.2159

cor(CherBoar$NDVI , CherBoar$Sr90) #-0.3418
cor(CherBoar$NDVI, CherBoar$Cs137) #-0.3322
cor(CherBoar$NDVI, CherBoar$Am241) #-0.3224
cor(CherBoar$NDVI, CherBoar$Pu238) #-0.3434
cor(CherBoar$NDVI, CherBoar$Pu239) #-0.3556
cor(CherBoar$NDVI, CherBoar$Pu240) #-0.3434

cor(CherBoar$Distance_power_plant, CherBoar$Sr90) #-0.5833
cor(CherBoar$Distance_power_plant, CherBoar$Cs137) #-0.4881
cor(CherBoar$Distance_power_plant, CherBoar$Am241) #-0.5671
cor(CherBoar$Distance_power_plant, CherBoar$Pu238) #-0.5862
cor(CherBoar$Distance_power_plant, CherBoar$Pu239) #-0.5823
cor(CherBoar$Distance_power_plant, CherBoar$Pu240) #-0.586

cor(CherBoar$Distance_closest_water, CherBoar$Sr90) #-0.1142
cor(CherBoar$Distance_closest_water, CherBoar$Cs137) #-0.0207
cor(CherBoar$Distance_closest_water, CherBoar$Am241) #-0.1197
cor(CherBoar$Distance_closest_water, CherBoar$Pu238) #-0.1188
cor(CherBoar$Distance_closest_water, CherBoar$Pu239) #-0.1286
cor(CherBoar$Distance_closest_water, CherBoar$Pu240) #-0.1188

cor(CherBoar$Distance_to_Road, CherBoar$Sr90) #-0.1894
cor(CherBoar$Distance_to_Road, CherBoar$Cs137) #-0.1735
cor(CherBoar$Distance_to_Road, CherBoar$Am241) #-0.1766
cor(CherBoar$Distance_to_Road, CherBoar$Pu238) #-0.1849
cor(CherBoar$Distance_to_Road, CherBoar$Pu239) #-0.1830
cor(CherBoar$Distance_to_Road, CherBoar$Pu240) #-0.18498

#radiation only
cor(CherBoar$Sr90, CherBoar$Cs137)
cor(CherBoar$Sr90, CherBoar$Am241)
cor(CherBoar$Sr90, CherBoar$Pu238)
cor(CherBoar$Sr90, CherBoar$Pu239)
cor(CherBoar$Sr90, CherBoar$Pu240)

cor(CherBoar$Cs137, CherBoar$Am241)
cor(CherBoar$Cs137, CherBoar$Pu238)
cor(CherBoar$Cs137, CherBoar$Pu239)
cor(CherBoar$Cs137, CherBoar$Pu240)

cor(CherBoar$Am241, CherBoar$Pu238)
cor(CherBoar$Am241, CherBoar$Pu239)
cor(CherBoar$Am241, CherBoar$Pu240)

cor(CherBoar$Pu238, CherBoar$Pu239)
cor(CherBoar$Pu238, CherBoar$Pu240)
cor(CherBoar$Pu238, CherBoar$Pu239) #0.9984546

cor(CherBoar$Pu239, CherBoar$Pu240)

CherBoarcovs <- CherBoar[,18:29]
M <- cor(CherBoarcovs)
library(corrplot)
corrplot(M, method = "number")

cor.test(CherBoarcovs$Sr90, CherBoarcovs$Am241)
cor.test(CherBoarcovs$Sr90, CherBoarcovs$Pu238)
cor.test(CherBoarcovs$Sr90, CherBoarcovs$Pu239)
cor.test(CherBoarcovs$Sr90, CherBoarcovs$Pu240)
cor.test(CherBoarcovs$Sr90, CherBoarcovs$Cs137)
cor.test(CherBoarcovs$Cs137, CherBoarcovs$Am241)
cor.test(CherBoarcovs$Cs137, CherBoarcovs$Pu238)
cor.test(CherBoarcovs$Cs137, CherBoarcovs$Pu239)
cor.test(CherBoarcovs$Cs137, CherBoarcovs$Pu240)
cor.test(CherBoarcovs$Am241, CherBoarcovs$Pu238)
cor.test(CherBoar$Am241, CherBoar$Pu239)
cor.test(CherBoar$Am241, CherBoar$Pu240)

cor.test(CherBoar$Pu238, CherBoar$Pu239)
cor.test(CherBoar$Pu238, CherBoar$Pu240)
cor.test(CherBoar$Pu238, CherBoar$Pu239)

cor.test(CherBoar$Pu239, CherBoar$Pu240)

cor.test(CherBoarcovs$Distance_power_plant, CherBoarcovs$Sr90)
cor.test(CherBoarcovs$Distance_power_plant, CherBoarcovs$Cs137)
cor.test(CherBoarcovs$Distance_power_plant, CherBoarcovs$Am241)
cor.test(CherBoarcovs$Distance_power_plant, CherBoarcovs$Pu238)
cor.test(CherBoarcovs$Distance_power_plant, CherBoarcovs$Pu239)
cor.test(CherBoarcovs$Distance_power_plant, CherBoarcovs$Pu240)

```

Single Covariate Models:
Will be testing a zero-inflated poisson (zip) and negative binomial (nb) distribution
```{r}
nb01 <- unmarked::pcount(~camdays ~ Lat., data = UMFcher, K = 858, mixture = "NB")
nb02 <- unmarked::pcount(~camdays ~ Long., data = UMFcher, K = 858, mixture = "NB")
nb03 <- unmarked::pcount(~camdays ~ Sr90, data = UMFcher, K = 858, mixture = "NB")
nb04 <- unmarked::pcount(~camdays ~ Cs137, data = UMFcher, K = 858, mixture = "NB")
nb05 <- unmarked::pcount(~camdays ~ Am241, data = UMFcher, K = 858, mixture = "NB")
nb06 <- unmarked::pcount(~camdays ~ Pu238, data = UMFcher, K = 858, mixture = "NB")
nb07 <- unmarked::pcount(~camdays ~ Pu239, data = UMFcher, K = 858, mixture = "NB")
nb08 <- unmarked::pcount(~camdays ~ Pu240, data = UMFcher, K = 858, mixture = "NB")
nb09 <- unmarked::pcount(~camdays ~ NDVI, data = UMFcher, K = 858, mixture = "NB")
nb010 <- unmarked::pcount(~camdays ~ Distance_to_Road, data = UMFcher, K = 858, mixture = "NB")
nb011 <- unmarked::pcount(~camdays ~ Distance_closest_water, data = UMFcher, K = 858, mixture = "NB")
nb012 <- unmarked::pcount(~camdays ~ Distance_power_plant, data = UMFcher, K = 858, mixture = "NB")

zip01 <- unmarked::pcount(~camdays ~ Lat., data = UMFcher, K = 858, mixture = "ZIP")
zip02 <- unmarked::pcount(~camdays ~ Long., data = UMFcher, K = 858, mixture = "ZIP")
zip03 <- unmarked::pcount(~camdays ~ Sr90, data = UMFcher, K = 858, mixture = "ZIP")
zip04 <- unmarked::pcount(~camdays ~ Cs137, data = UMFcher, K = 858, mixture = "ZIP")
zip05 <- unmarked::pcount(~camdays ~ Am241, data = UMFcher, K = 858, mixture = "ZIP")
zip06 <- unmarked::pcount(~camdays ~ Pu238, data = UMFcher, K = 858, mixture = "ZIP")
zip07 <- unmarked::pcount(~camdays ~ Pu239, data = UMFcher, K = 858, mixture = "ZIP")
zip08 <- unmarked::pcount(~camdays ~ Pu240, data = UMFcher, K = 858, mixture = "ZIP")
zip09 <- unmarked::pcount(~camdays ~ NDVI, data = UMFcher, K = 858, mixture = "ZIP")
zip010 <- unmarked::pcount(~camdays ~ Distance_to_Road, data = UMFcher, K = 858, mixture = "ZIP")
zip011 <- unmarked::pcount(~camdays ~ Distance_closest_water, data = UMFcher, K = 858, mixture = "ZIP")
zip012 <- unmarked::pcount(~camdays ~ Distance_power_plant, data = UMFcher, K = 858, mixture = "ZIP")

#Testing model fit and returning AIC values
fitest1 <- unmarked::fitList(nb01, nb02, nb03, nb04, nb05, nb06, nb07, nb08, nb09, nb010, nb011, nb012, zip01, zip02, zip03, zip04, zip05, zip06, zip07, zip08, zip09, zip010, zip011, zip012)
testset1 <- unmarked::modSel(fitest1) 
AICtestset1 <- data.frame(testset1@Full[["model"]],testset1@Full[["AIC"]])
#negative binomial models for each covariate return the lowest AIC, so all future models will be made with the negative binomial mixture
testset1
#so far, model nb02 has the lowest overall AIC - 4268.44
```


Corrected AIC values:
```{r}
MuMIn::AICc(nb01, nb02, nb03, nb04, nb05, nb06, nb07, nb08, nb09, nb010, nb011, nb012)

newlist <- list(null, nb01, nb02, nb03, nb04, nb05, nb06, nb07, nb08, nb09, nb010, nb011, nb012)
AICcmodavg::aictab(newlist)
```

Model Set 2: Additive Models
All combinations of Long. with other covariates
```{r}

fm.1 <- unmarked::pcount(~camdays ~ Lat. + Long., data = UMFcher, K = 858, mixture = "NB")
fm.2 <- unmarked::pcount(~camdays ~ Long. + Sr90, data = UMFcher, K = 858, mixture = "NB")
fm.3 <- unmarked::pcount(~camdays ~ Long. + Am241, data = UMFcher, K = 858, mixture = "NB")
fm.4 <- unmarked::pcount(~camdays ~ Long. + Pu238, data = UMFcher, K = 858, mixture = "NB")
fm.5 <- unmarked::pcount(~camdays ~ Long. + Pu239, data = UMFcher, K = 858, mixture = "NB")
fm.6 <- unmarked::pcount(~camdays ~ Long. + Pu240, data = UMFcher, K = 858, mixture = "NB")
fm.7 <- unmarked::pcount(~camdays ~ Long. + NDVI, data = UMFcher, K = 858, mixture = "NB")
fm.8 <- unmarked::pcount(~camdays ~ Long. + NDVI, data = UMFcher, K = 858, mixture = "NB")
fm.9 <- unmarked::pcount(~camdays ~ Long. + Distance_to_Road, data = UMFcher, K = 858, mixture = "NB")
fm.10 <- unmarked::pcount(~camdays ~ Long. + Distance_closest_water, data = UMFcher, K = 858, mixture = "NB")
fm.11 <- unmarked::pcount(~camdays ~ Long. + Distance_power_plant, data = UMFcher, K = 858, mixture = "NB")

fl <- unmarked::fitList(fm.1, fm.2, fm.3, fm.4, fm.5, fm.6, fm.7, fm.8, fm.9, fm.10, fm.11)
AICtestset2 <- unmarked::modSel(fl)
AICtestset2
```

Corrected AICc values
```{r}
AICcs <- MuMIn::AICc(fm.1, fm.2, fm.3, fm.4, fm.5, fm.6, fm.7, fm.8, fm.9, fm.10, fm.11)
AICcs
#model fm.1 has the lowest corrected AIC from round 2 (4270.185) but it is not lower than that of nb.02 from the first round

newlist2 <- list(null, nb02, fm.1, fm.2, fm.3, fm.4, fm.5, fm.6, fm.7, fm.8, fm.9, fm.10, fm.11)
AICcmodavg::aictab(newlist2, modnames = c("null", "nb02", "fm.1", "fm.2", "fm.3", "fm.4", "fm.5", "fm.6", "fm.7", "fm.8", "fm.9", "fm.10", "fm.11"))

vif(mod = fm.1, type = "state")
vif(mod = fm.4, type = "state")
vif(mod = fm.3, type = "state")
```

Just trying 3
```{r}
threecovs <- unmarked::pcount(~camdays ~ Lat. + Long. + Pu238, data = UMFcher, K = 858, mixture = "NB")
#AICc is higher than others
vif(mod = threecovs, type = "state")
```


RAI Barplot per Year
```{r}
CherRAIs <- read_excel("Sophia Thesis- Cher ALL.xlsx", sheet = "RAI per year")
M2016 <- mean(na.omit(CherRAIs$RAI2016))
M2017 <- mean(na.omit(CherRAIs$RAI2017))
M2018 <- mean(na.omit(CherRAIs$RAI2018))
M2019 <- mean(na.omit(CherRAIs$RAI2019))
M2020 <- mean(na.omit(CherRAIs$RAI2020))
M2021 <- mean(na.omit(CherRAIs$RAI2021))
M2022 <- mean(na.omit(CherRAIs$RAI2022))
M2023 <- mean(na.omit(CherRAIs$RAI2023))

V2016 <- var(na.omit(CherRAIs$RAI2016))
V2017 <- var(na.omit(CherRAIs$RAI2017))
V2018 <- var(na.omit(CherRAIs$RAI2018))
V2019 <- var(na.omit(CherRAIs$RAI2019))
V2020 <- 0
V2021 <- var(na.omit(CherRAIs$RAI2021))
V2022 <- var(na.omit(CherRAIs$RAI2022))
V2023 <- var(na.omit(CherRAIs$RAI2023))

newframe <- data.frame(Year=c('2016', '2017', '2018', '2019', '2020', '2021', '2022', '2023'), value=c(M2016, M2017, M2018, M2019, M2020, M2021, M2022, M2023), varr=c(V2016, V2017, V2018, V2019, V2020, V2021, V2022, V2023))

ggplot(data = newframe, aes(x=Year, y=value)) + geom_bar(stat = "identity", fill =  c("#2A788EFF", "#75D054FF", "#FDE725FF", "#481567FF", "lightseagreen", "darkseagreen2", "darkgoldenrod2", "darkorchid2")) + ylab("Mean RAI Value") + ylim(-0.025,0.35) + geom_errorbar(aes(x=Year, ymin=value-varr, ymax=value+varr), width = 0.2, color = "black", alpha = 0.9) + theme_bw(base_size = 15)

```

Boxplot instead?
```{r}
boxframe16 <- data.frame(RAI = na.omit(CherRAIs$RAI2016), Year = rep(2016, 46))
boxframe17 <- data.frame(RAI = na.omit(CherRAIs$RAI2017), Year = rep(2017, 42))
boxframe18 <- data.frame(RAI = na.omit(CherRAIs$RAI2018), Year = rep(2018, 52))
boxframe19 <- data.frame(RAI = na.omit(CherRAIs$RAI2019), Year = rep(2019, 64))
boxframe20 <- data.frame(RAI = na.omit(CherRAIs$RAI2020), Year = rep(2020, 1))
boxframe21 <- data.frame(RAI = na.omit(CherRAIs$RAI2021), Year = rep(2021, 19))
boxframe22 <- data.frame(RAI = na.omit(CherRAIs$RAI2022), Year = rep(2022, 28))
boxframe23 <- data.frame(RAI = na.omit(CherRAIs$RAI2023), Year = rep(2023, 23))

boxframe <- rbind(boxframe16, boxframe17, boxframe18, boxframe19, boxframe20, boxframe21, boxframe22, boxframe23)

boxframe$Year <- as.factor(boxframe$Year)

ggplot(data = boxframe, aes(x=Year, y=RAI)) + geom_boxplot(fill = c("#2A788EFF", "#75D054FF", "#FDE725FF", "#481567FF", "lightseagreen", "darkseagreen2", "darkgoldenrod2", "darkorchid2")) + coord_flip(ylim=c(0,1.5)) + theme_bw(base_size=15)
```

Plotting Model Predicted Values:
```{r}
nb02preds <- predict(nb02, type = "state")
#model gives counts, not RAI

totalCount <- read_excel("Sophia Thesis- Cher ALL.xlsx", sheet = "Counts")

predsANDcounts <- cbind(nb02preds, totalCount)


plot(nb02preds$Predicted ~ predsANDcounts$Long., ylab="Raw Abundance", xlab="Longitude (°E)", ylim=c(0,400), pch=17)
points(x = predsANDcounts$Long., y = predsANDcounts$Total_obs, col = "forestgreen", pch=17)
legend(x = 29.5, y = 350, legend = c("Model-Predicted Values", "Recorded Counts"), col = c("black", "forestgreen"), pch = 17)

##OR, convert model predicted values to RAI
predsANDcounts$modelRAI <- predsANDcounts$Predicted/predsANDcounts$total_days

plot(predsANDcounts$modelRAI ~ predsANDcounts$Long., ylab="RAI", xlab="Longitude (°E)", ylim = c(0,6), pch=17)
points(x = predsANDcounts$Long., y = predsANDcounts$RAI, col = "forestgreen", pch=17)
legend(x = 29.5, y = 5, legend = c("Model-Predicted RAI", "Observed RAI"), col = c("black", "forestgreen"), pch = 17)
```

```{r}
#Log(x+1) transforming RAI
boxframe16$transform <- log(boxframe16$RAI + 1)
boxframe17$transform <- log(boxframe17$RAI + 1)
boxframe18$transform <- log(boxframe18$RAI + 1)
boxframe19$transform <- log(boxframe19$RAI + 1)
boxframe20$transform <- log(boxframe20$RAI + 1)
boxframe21$transform <- log(boxframe21$RAI + 1)
boxframe22$transform <- log(boxframe22$RAI + 1)
boxframe23$transform <- log(boxframe23$RAI + 1)

boxframe <- rbind(boxframe16, boxframe17, boxframe18, boxframe19, boxframe20, boxframe21, boxframe22, boxframe23)
boxframe$Year <- as.factor(boxframe$Year)

ggplot(data = boxframe, mapping = aes(x=Year, y=transform)) + geom_boxplot(fill = c("#2A788EFF", "#75D054FF", "#FDE725FF", "#481567FF", "lightseagreen", "darkseagreen2", "darkgoldenrod2", "darkorchid2")) + labs(y = "Log(RAI + 1)", title = "Chernobyl") + theme_bw(base_size = 15) + ylim(0,0.5)

```
